\section{Related Work}

\todo[inline]{Trim and compress the related work section}

\subsection{Distributed storage and synchronization systems}

\towrite{Related Works: Git}
\towrite{Related Works: Mercurial}

\paragraph{Camlistore}

Camlistore~\cite{camlistore_homepage} is an open-source project to create a private long-term data storage system for personal users.
It allows storage of diverse types of data and it synchronizes between multiple replicas of the data store.
However, it eschews normal filesystems and creates its own schemas to store various media.


\paragraph{Dat Data}

Dat~\cite{dat_homepage} is an open-source project for publishing and sharing scientific data sets for research.
This project has a lot of overlap with ours, and several of the core ideas are similar, including breaking files into smaller chunks, and tracking changes via a Git-like \gls{DAG}.
However, their focus is different.
The Dat team is concentrating on publishing research data, and making that specific task as simple as possible for non-technical researchers who might not be familiar with version control.
By contrast, our project operates at a lower level of abstraction, offering the full power of version control in a very general way, exposing and illuminating the complexities rather than trying to hide them or automate them away.

Where Dat focuses on publishing on the open internet, we focus on ad-hoc networks and data that may be private.
Where Dat has components for automating peer discovery and consensus, we work at a lower level, trying to perfect and generalize the storage aspect first.
Dat seems to assume that data sets will be small enough to fit on a typical disk on a workstation, while we want to scale even larger.

We hope that our system could be used as a base to build something like Dat, but we intend for \gls{DMV} to be more general than the Dat core.


\paragraph{Eyo}

Eyo~\cite{Strauss:2011:EDP:2002181.2002216} is system for storing personal media and synchronizing it between devices.
It utilizes a Git-like content-addressed \gls{objectstore} behind the scenes, but it works more like a networked filesystem than version control.
It focuses on organizing media by metadata, which requires agreement on metadata formats, and it requires applications to be rewritten to access files via Eyo rather than the filesystem, both of which are thorny and ambitious problems.
We focus purely on storage and synchronization.


\paragraph{Git-Annex and Git-Media}

Git-annex~\cite{git_annex_homepage} and Git-media~\cite{git_media_github} are open-source projects that extend Git with special handling for larger files.
Both store information about the larger files in the normal Git \gls{repository} and then store the files themselves in a separate location.
Git-media stores all the larger files in a separate data store which may be remote.
Git-annex is more flexible.
Annex files may be spread across several different remote \gls{repository} clones or data stores, and Git-annex has features for tracking the locations of annex files in different remote \glspl{repository} and moving them from one \gls{repository} to another.
These tracking and distribution features are very similar to our goals.
However, Git-annex is not quite as flexible as we aim for with \gls{DMV}.
It considers the large files atomic units, and it does not break them into smaller chunks for de-duplication.
Also, because metadata is processed by Git, it has the same limitations that Git does.
All \glspl{repository} must have all metadata, and performance suffers when metadata is too large to fit into RAM.


\paragraph{IPFS: The Interplanetary Filesystem}

IPFS~\cite{ipfs_github_main} is an open-source pro\-ject to create a global content-addressed filesystem.
By its global nature, all files are stored publicly, in a global network of nodes with global addressing.
IPFS should be an excellent resource for storing published information, but we want \gls{DMV} to work with private data sets.
We want individuals and organizations to be able manage their own data stores privately on their own hardware.

It should be noted that IPFS does have support for storing private objects by way of object-level encryption.
However, this seems wasteful of disk space, since small changes in the plain text of a file would completely change the ciphertext, leaving no way to compress the redundancy.


\paragraph{Kademlia}

Kademlia~\cite{Maymounkov2002} is an advanced distributed hash table system that updates its network topology information as part of normal lookups.
Its focus is on efficient routing between nodes, while we are focusing first on storage.


\paragraph{Rsync}
\label{related-rsync}

Rsync~\cite{rsynctechreport} is a utility that synchronizes files across a low-band\-width network link by using a rolling hash to find similar parts of the file, and then transfer only those portions of the file that have changed.
Its \glsfirst{rollinghash} is the inspiration for the chunk-splitting algorithms used by Rsyncrypto~\cite{rsyncrypto_algorithm}, Bup~\cite{bup_design}, and \gls{DMV} (\autoref{chunking-algorithm}).
As a utility, Rsync focuses exclusively on synchronizing files.
It does not track history or the relation between files.

Rsync coauthor Andrew Tridgell is also indirectly responsible for the development of Git.
His work to reverse-engineer the BitKeeper protocol is what caused BitMover to revoke the special free BitKeeper license for Linux kernel developers, which is what led Linus Torvalds to develop his own \gls{SCM}~\cite{git_10_years_interview}.


\paragraph{Rsyncrypto}
\label{related-rsyncrypto}

Rsyncrypto~\cite{rsyncrypto_algorithm} is an encryption system that uses a roll\-ing hash to encrypt files by chunks, so that the encrypted files are still "rsync-able." Normally, a goal of an encryption algorithm is to have small changes in the input lead to a completely different ciphertext.
This is good for security, but it negates the advantages of rsync --- if the entire file is different, there are no advantages to be had in transferring only the changed parts when synchronizing files.
Rsync uses a rolling hash to break the file into chunks and encrypt the chunks separately, so that a small change in the input will only change the ciphertext for the chunk that contains it, and so rsync can transfer only the chunk that changes.

Rsyncrypto's uses the same \glsfirst{rollinghash} as Rsync~\cite{rsyncrypto_algorithm,rsynctechreport}, setting chunk boundaries where the checksum is zero.
\gls{DMV} uses this same chunking algorithm (see \autoref{chunking-algorithm}).


\subsection{De-duplicating Storage and Backup}

\paragraph{Boar}

Boar~\cite{boar_homepage} is an open-source project to create a version control system for large binary files.
It is one of the main inspirations for our project.
It stores file versions in a content-addressed way, and provides de-duplication for large files that only change in small pieces, and it can truncate history to reclaim disk space.
However, Boar retreats to the centralized version control paradigm, with a central \gls{repository} that working directories must connect to in order to check files in or out.
We want to provide the advantages of Boar in a flexible distributed version control model.
Boar also has practical limitations on \gls{repository} size and number of files.
\glspl{repository} are assumed to fit on one disk volume, and file metadata is assumed to fit into Ram.
\gls{DMV} tries to avoid both of those limitations.


\paragraph{Bup}\label{related_bup}

Bup~\cite{bup_homepage} is an open-source file backup system that is based on Git's \gls{repository} format.
It does many of the things that \gls{DMV} does: it breaks files into chunks by \gls{rollinghash}, and it has considerations for metadata that is larger than RAM.
Bup is one of the systems we evaluated along with Git and Mercurial (\autoref{file-size-exp-desc}), and we determined that Bup's manner of storing chunks is ideal for the next step for \gls{DMV} (\autoref{chunk-then-recombine}).
However, though Bup is built on \glsdisp{VCS}{version control}, it is locked into a backup-based workflow.
History is linear and based on clock time of backup.
And it assumes that the whole data set and the whole \gls{repository} can fit onto one filesystem.
\gls{DMV} is built to be more general and versatile.


\paragraph{Time Machine}

Time Machine~\cite{timemachine_patent} is a backup utility from Apple, included in Mac OS X Leopard, that makes frequent automated file hierarchy backups, using filesystem hard links to de-duplicate unchanged files~\cite{timemachine_magic}. This de-duplication allows it to store many different backup versions. Time Machine's functionality can be mimicked by using Rsync with the \lstinline{--link-dest} option, which also hard-links unchanged files during a recursive sync~\cite{timemachine_foreveryunix}.
