\chapter{Conclusion and Summary of Contributions}

\glsreset{DMV}

In this dissertation, we have examined the cryptographic \acrfull{DAG} as a data
structure for data storage, and the ways that it can be sliced to shard data
across nodes in a distributed system, according to what data is needed locally
at each location.

We have performed experiments to probe the scalability limits of existing
\gls{DAG}-based \acrlongpl{DVCS}. We have shown that the maximum size of file
that existing systems like Git and Mercurial can store is limited by the amount
of available memory in the system, and we conclude that it is because those
systems calculate deltas of files to de-duplicate data, and they load the entire
file into memory in order to do so.

We have also rediscovered the limits of the Unix filesystem for storing many
small files.

\begin{itemize}

    \item Writing files smaller than the filesystem block size incurs storage
        overhead.

    \item Splitting files among too many subdirectories takes \gls{inode} that
        are needed to store files.

    \item Jumping between directories when writing files incurs write-speed
        penalties.

\end{itemize}

We have shown that any \gls{VCS} that stores objects as individual files on the
filesystem will encounter these filesystem limitations as they try to scale in
terms of number of files. \glspl{VCS} that also break files into chunks will
turn the problem of storing large files into the problem of storing many files,
and will also encounter these limitations. However, the limitations can be
avoided by aggregating objects into \glspl{packfile} as Bup does.

And finally, we have described the idea, architecture, design, and
implementation of a distributed data storage system we call \gls{DMV} that
expands on the \glsdisp{DVCS}{distributed version control} concept to store
larger and more diverse data sets, with a high degree of control over data
locality, and an availability to write updates for any data held locally. Though
time constraints prevented us from implementing the network features we had
planned, the \gls{DMV} prototype has enough functionality to be experimented on
against existing \acrlongpl{DVCS} and to demonstrate the addition of new
\glspl{commit} to a partial \gls{DAG}.
