\chapter{Abstract}

\glsunsetall
\glsreset{DMV}

%\perottoinline{1. What's wrong with the world?}

It is still strangely difficult to backup and synchronize data. Cloud computing
solves the problem by centralizing everything and letting someone else handle
the backups. But what about situations with low connectivity or sensitive data?

%\perottoinline{2. Motivation (why bother?)}

For this, software developers have an interesting distributed, decentralized,
and partition-tolerant data storage system right at their fingertips:
\glsdisp{DVCS}{distributed version control}.

%\perottoinline{3. How we did better: idea, architecture, design, implementation.}

Inspired by \glsdisp{DVCS}{distributed version control}, we have researched and
developed a prototype for a scalable high-availability system called \gls{DMV}.
\gls{DMV} expands Git's data model  to allow files to be broken into more
digestible chunks via a \gls{rollinghash}. \gls{DMV} will also allow data to be
sharded according to data locality needs, slicing the data set in space (subset
of data with full history), time (subset of history for full data set), or both.
\gls{DMV} \glspl{repository} will be able to read and to update any subset of
the data that they have locally, and then synchronize with other
\glspl{repository} in an ad-hoc network.

%\perottoinline{4. Experiments; results}

We have performed experiments to probe the scalability limits of existing
version control systems, specifically what happens as file sizes grow ever
larger or as the number of files grow. We found that processing files whole
limits maximum file size to what can fit in RAM, and that storing millions of
objects loose as files with hash-based names incurs disk space overhead and
write speed penalties. We have observed a system needing \SI{24}{\second} to
store a \SI{6.8}{\kib} file.

%\perottoinline{5. Lessons learned}



%\perottoinline{6. Main conclusions}

We conclude that the key to storing large files is the break them into many
small chunks, and that the key to storing many chunks is to aggregate them into
pack files. And though the current \gls{DMV} prototype does only the former, we
have a clear path forward as we continue our work.

%


\glsresetall
